Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-24 paper.tex
fixed bibliography style
small fixes

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace},\footnote{For our domain we do not require the trace to be ordered -- so it is similar to, but not exactly the same as, the execution trace of the program.} since it corresponds to the commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them.
For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the input. The network is trained from an automatically-generated corpus of synthetic image-generating programs.

The second step involves generating a high-level program capable of producing the trace identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
This paper argues that this stage is best achieved by \emph{constraint-based program synthesis} \cite{solar2008program}. The program synthesizer can search the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, repetition, or conditional branches.
Although these program synthesizers do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\cite{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[h]
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes drawing command inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\end{table}

@@

\setlength\tabcolsep{3.5pt}
\begin{figure}[H]
  \begin{minipage}[t]{8.25cm}
    \begin{tabular}{llll}
      \includegraphics[width = 2cm]{figures/expert-38.png}&
      \includegraphics[width = 2cm]{figures/expert-73.png}&
      \includegraphics[width = 2cm]{figures/expert-77.png}&
      \includegraphics[width = 2cm]{figures/expert-21.png}
      \\
      \includegraphics[width = 2cm]{figures/38-parse.png}     &
      \includegraphics[width = 2cm]{figures/73-parse.png}     &
      \includegraphics[width = 2cm]{figures/77-parse.png}    &
      \includegraphics[width = 2cm]{figures/21-parse.png}    
      \end{tabular}
    \subcaption{Left to right: Ising model, recurrent network architecture, figure from a deep learning textbook (Goodfellow et al:~\cite{Goodfellow-et-al-2016}), graphical model}
  \end{minipage}
\hfill  \begin{minipage}[t]{4.25cm}
    \begin{tabular}{ll}
          \includegraphics[width = 2cm]{figures/expert-34.png}&
          \includegraphics[width = 2cm]{figures/expert-1.png}\\
              \includegraphics[width = 2cm]{figures/34-parse.png}     &
      \includegraphics[width = 2cm]{figures/expert-1-parse.png}    
    \end{tabular}

@@

\begin{table}[H]
  \begin{tabular}{rl}\toprule
  Program$\to$&Statement; $\cdots$; Statement\\
  Statement$\to$&\texttt{circle}(Expression,Expression)\\
  Statement$\to$&\texttt{rectangle}(Expression,Expression,Expression,Expression)\\
  Statement$\to$&\texttt{line}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Statement$\to$&\texttt{for}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ if }$(\text{Var} > 0)$\texttt{ \{ }Program\texttt{ \}; }Program\texttt{ \}}\\
  Statement$\to$&\texttt{reflect}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&$\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$\\
  Var$\to$&A free (unused) variable\\
  $\mathbb{Z}$\to$&an integer\\
  Axis$\to$&\texttt{X = }$\mathbb{Z}$ | \texttt{Y = }$\mathbb{Z}$\\\bottomrule
  \end{tabular}
  \caption{Grammar over graphics programs. We allow loops (\texttt{for}) with conditionals (\texttt{if}), vertibb/horizontal reflections (\texttt{reflect}), variables (Var) and affine transformations ($\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$).}\label{DSL}
\end{table}

@@

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see~\cite{solar2008program}.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
In practice, synthesis times vary from minutes to hours,
with 16\% of the diagrams timing out the synthesizer after 3 hours.
See Figure~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving most problems in under a minute.

@@

\begin{table}[h]\centering
  \begin{tabular}{lll}\toprule
  Parameter&Description&Range\\\midrule
  Loops?&Is the program allowed to loop?&$\{\text{True},\text{False}\}$\\
  Reflects?&Is the program allowed to have reflections?&$\{\text{True},\text{False}\}$\\
  Incremental?&Solve the problem piece-by-piece or all at once?&$\{\text{True},\text{False}\}$\\
  Maximum depth& Bound on the depth of the program syntax tree&$\{1,2,3\}$
  \\\bottomrule
  \end{tabular}
  \caption{Parameterization of different ways of posing the program synthesis problem. The policy learns to choose parameters likely to quickly yield a minimal cost program. In our notation, we write $\sigma $ to mean an assignment to each of these parameters, so $\sigma $ can assume one of 24 different values.}\label{policyOutput}
  \end{table}

Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm} (see Schmidhuber 2004:~\cite{schmidhuber2004optimal}):

@@

\bibliographystyle{unsrt} {\small \bibliography{main}}

>>>>

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace},\footnote{For our domain we do not require the trace to be ordered -- so it is similar to, but not exactly the same as, the execution trace of the program.} since it corresponds to the commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them.
For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.

The second step involves generating a high-level program capable of producing the trace identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \cite{solar2008program}. The program synthesizer can search the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, repetition, or conditional branches.
Although these program do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\cite{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[h]
  \caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes drawing command inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\end{table}

@@

\setlength\tabcolsep{3.5pt}
\begin{figure}[H]
  \begin{minipage}[t]{8.25cm}
    \begin{tabular}{llll}
      \includegraphics[width = 2cm]{figures/expert-38.png}&
      \includegraphics[width = 2cm]{figures/expert-73.png}&
      \includegraphics[width = 2cm]{figures/expert-77.png}&
      \includegraphics[width = 2cm]{figures/expert-21.png}
      \\
      \includegraphics[width = 2cm]{figures/38-parse.png}     &
      \includegraphics[width = 2cm]{figures/73-parse.png}     &
      \includegraphics[width = 2cm]{figures/77-parse.png}    &
      \includegraphics[width = 2cm]{figures/21-parse.png}    
      \end{tabular}
    \subcaption{Left to right: Ising model, recurrent network architecture, figure from a deep learning textbook \cite{Goodfellow-et-al-2016}, graphical model}
  \end{minipage}
\hfill  \begin{minipage}[t]{4.25cm}
    \begin{tabular}{ll}
          \includegraphics[width = 2cm]{figures/expert-1.png}&
          \includegraphics[width = 2cm]{figures/expert-34.png}\\
      \includegraphics[width = 2cm]{figures/expert-1-parse.png}     &
              \includegraphics[width = 2cm]{figures/34-parse.png}    
    \end{tabular}

@@

\begin{table}[H]
  \caption{Grammar over graphics programs. We allow loops (\texttt{for}) with conditionals (\texttt{if}), vertibb/horizontal reflections (\texttt{reflect}), variables (Var) and affine transformations ($\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$).}\label{DSL}
  \begin{tabular}{rl}\toprule
  Program$\to$&Statement; $\cdots$; Statement\\
  Statement$\to$&\texttt{circle}(Expression,Expression)\\
  Statement$\to$&\texttt{rectangle}(Expression,Expression,Expression,Expression)\\
  Statement$\to$&\texttt{line}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Statement$\to$&\texttt{for}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ if }$(\text{Var} > 0)$\texttt{ \{ }Program\texttt{ \}; }Program\texttt{ \}}\\
  Statement$\to$&\texttt{reflect}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&$\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$\\
  Var$\to$&A free (unused) variable\\
  $\mathbb{Z}\to$&an integer\\
  Axis$\to$&\texttt{X = }$\mathbb{Z}$ | \texttt{Y = }$\mathbb{Z}$\\\bottomrule
  \end{tabular}
\end{table}

@@

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
In practice, synthesis times vary from minutes to hours,
with 16\% of the diagrams timing out the synthesizer after 3 hours.
See Figure~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving most problems in under a minute.

@@

\begin{table}[h]\centering
  \caption{Parameterization of different ways of posing the program synthesis problem. The policy learns to choose parameters likely to quickly yield a minimal cost program. In our notation, we write $\sigma $ to mean an assignment to each of these parameters, so $\sigma $ can assume one of 24 different values.}\label{policyOutput}
  \begin{tabular}{lll}\toprule
  Parameter&Description&Range\\\midrule
  Loops?&Is the program allowed to loop?&$\{\text{True},\text{False}\}$\\
  Reflects?&Is the program allowed to have reflections?&$\{\text{True},\text{False}\}$\\
  Incremental?&Solve the problem piece-by-piece or all at once?&$\{\text{True},\text{False}\}$\\
  Maximum depth& Bound on the depth of the program syntax tree&$\{1,2,3\}$
  \\\bottomrule
  \end{tabular}
  \end{table}

Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm} (see~\cite{schmidhuber2004optimal}):

@@

\bibliographystyle{iclr2018_conference}
                  {\small \bibliography{main}}
